### 2018-05-24 InequalitiesPaper/Inequality3.tex
Adding all research and expository material so far

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,geometry,graphicx,enumerate}

\def\Name{Edward Kim}  % Your name

%for code%
\usepackage{listings}
\lstset{language=python}

\title{Proof of Inequality 2}
\author{\Name}

\textheight=9in
\textwidth=6.5in
\topmargin=-.75in
\oddsidemargin=0.25in
\evensidemargin=0.25in
\DeclareMathSizes{11}{19}{13}{9}
\setlength\parindent{0pt}

\begin{document}
\maketitle

First, let us state the problem and develop some motivation for it: \newline

Let $n \in \mathbb{N}$. Then the following holds provided that $a_i \geq 1$ for $1 \leq i \leq n$.

\begin{equation}
 \sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \frac{n}{1 + (a_1...a_n)^{\frac{1}{n}}}
\end{equation}

\textbf{Proposition: } The inequality holds for all $n \in  \mathbb{N}$. \newline

\textbf{Proof: } We shall use the property that the inequality holds for all $n = 2^k$ for some $k \in \mathbb{N}$.

Let $m$ be a power of 2 such that $m > n$. Observe the following modification to the original sum as follows.

\begin{gather}
  \sum_{i=1}^{n} \frac{1}{1 + a_i} + \sum^{m-n} \frac{1}{1 + (a_1...a_n)^{\frac{1}{n}}} 
\end{gather}

Since $(a_1...a_n)^{\frac{1}{n}} \geq 1$ and there are $m$ terms in the sum, we can apply the inequality:

\begin{gather}
\sum_{i=1}^{n} \frac{1}{1 + a_i} + \frac{m-n}{1 + (a_1...a_n)^{\frac{1}{n}}} \geq \frac{m}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} \\ 
\sum_{i=1}^{n} \frac{1}{1 + a_i} + \frac{m-n}{1 + (a_1...a_n)^{\frac{1}{n}}} \geq \frac{n}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} + \frac{m-n}{1 + ((a_1...a_n)^{\frac{m}{n}})^{\frac{1}{m}}} \\
\sum_{i=1}^{n} \frac{1}{1 + a_i} \geq \frac{n}{1 + (a_1...a_n)^{\frac{1}{n}}}
\end{gather}

Thus, the inequality holds for any $n \in \mathbb{N}$.

\end{document}