### 2021-01-26 misc/InequalitiesPaper/Inequality2.tex

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,geometry,graphicx,enumerate}

\def\Name{Edward Kim}  % Your name

%for code%
\usepackage{listings}
\lstset{language=python}

\title{Proof of Inequality 2}
\author{\Name}

\textheight=9in
\textwidth=6.5in
\topmargin=-.75in
\oddsidemargin=0.25in
\evensidemargin=0.25in
\DeclareMathSizes{11}{19}{13}{9}
\setlength\parindent{0pt}

\begin{document}
 \maketitle

 (Holy crap. This induction took me hours. I was trying all these weird methods when the simplest answer was right in front of me.)
 \newline \newline
 \noindent

 Let us first prove the base case for when $n = 2$. This is proven in Inequality 1 so the base case is proven. Assume
 that the inequality holds for some $k$ so that $n = 2^k$. Now let us prove the inequality for 
 $2^{k+1}$. \newline

 \noindent

 We have the sum:

 \begin{equation}
  \sum_{i=1}^{2^{k+1}} \frac{1}{1+a_i}.
 \end{equation}

 Now choose any two partitions and relabel the $a_i$ such that the sum becomes:

  \begin{equation}
  \sum_{i=1}^{2^{k+1}} \frac{1}{1+a_i} = \sum_{i=1}^{2^{k}} \frac{1}{1+a_i} + \sum_{i=1}^{2^{k}} \frac{1}{1+\widetilde{a_i}}.
 \end{equation}

 Let us denote the sum as $\phi$. \newline

 By our induction hypothesis,

 \begin{equation}
  \phi \geq \frac{2^k}{1 + (a_1...a_{2^k})^{\frac{1}{2^k}}} + \frac{2^k}{1 + (\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}}}
 \end{equation}

Let the above sum be denoted as $R$. Factor out the $2^k$

\begin{equation}
 \phi \geq 2^k(\frac{1}{1 + (a_1...a_{2^k})^{\frac{1}{2^k}}} + \frac{1}{1 + (\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}}})
\end{equation}

However, $(a_1...a_{2^k})^{\frac{1}{2^k}} \geq 1$ and $(\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}} \geq 1$. So we can
use our induction hypothesis again!

\begin{equation}
 \phi \geq R \geq 2^k(\frac{2}{1 + \sqrt{((a_1...a_{2^{k}})^{\frac{1}{2^k}})((\widetilde{a_1}...\widetilde{a_{2^k}})^{\frac{1}{2^k}})}})
\end{equation}

However, this can be simplified to

\begin{gather}
 \phi \geq 2^k(\frac{2}{1 + \sqrt{(a_1...a_{2^{k+1}})^{\frac{1}{2^{k}}}}}) \\
 \phi \geq \frac{2^{k+1}}{1 + (a_1...a_{2^{k+1}})^{\frac{1}{2^{k+1}}}}
\end{gather}

This proves the induction and, thus, we have proved the inequality.
\end{document}