### 2021-11-17 Complexity/UGC/ugcmaxcut.tex

\section{UG-hardness of MAXCUT}
\subsection{Intuitions}
\subsection{Goemans-Williams Algorithm for MAXCUT}