### 2021-11-18 Complexity/UGC/hastad.tex
First draft of UGC paper

\section{H\aa stad's 3-bit PCP}
%Arora, Barak
\subsection{Aside on Dictatorship Testing}
To be written...
\subsection{Long Code}
To be written...