### 2022-07-02 thesis/frontmatter/acknowledgement.tex
Prelim thesis commit

%Acknowledgements are the author's statement of gratitude to and
%recognition of the people and institutions who helped the author's
%research and writing.

\begin{center}
\vspace*{52pt}
{\Large \textbf{ACKNOWLEDGEMENTS}}
\end{center}

I would like to thank the UNC Computer Science community for supporting my work and providing me a warm community for pursuing this research. In particular, I would like to thank my advisor, Parasara Sridhar Duggirala, for the opporunity to perform research and participate in his group.
%
%His choice to give me more autonomy to take classes at Duke University really helped me understand my academic interests better.
%
Bineet Ghosh, Manish Goyal, Abel Karimi, and Meghan Stuart were all great colleagues and I enjoyed my time exchanging ideas, proofreading drafts, and socializing over group lunches. I would also like to thank Stanley Bak for helping me improve the implementation of Kaa and providing valuable feedback during this research. Juan Garcia made a wonderful collegue and an equally wonderful friend. The nights we raided Peabody and Phillips hall was certainly a memorable time. Finally, I would like to thank my parents and my family for their ardent, un-ending support.

\clearpage