### 2022-07-02 thesis/results.tex

\chapter{Experimental Results}

\subsection{Kaa}
\label{sec:kaa}

\subsection{Benchmarks}
\label{sec:benchmarks}

\subsection{COVID19 Supermodel}
\label{sec:covid}

\subsection{Comparison of Template Generation Techniques}
\label{sec:compare}