### 2022-07-02 thesis/prelim.tex
Static Algorithm section started

\chapter{Preliminaries}
\label{chap:prelim}

We begin with some basic definitions pertaining to reachability and parallelotopes. The definition of Bernstein polynomials and the reachable set computaion algorithm will be defined. Finally, an outline of the reachability algorithm given by \cite{dreossi2016parallelotope} for polynomial dynamical systems will be presented.

\section{Basic Definitions}
\label{sec:definitions}

As stated in the previous sections, this thesis pertains to the reachability analysis of dynamical systems. Roughly speaking, a dynamical system is governed by a set of differential equations such that the states of the system evolve according the solutions of the said differential equations.
%
The state of a system, denoted as $x$, lies in a domain $D \subseteq \reals^n$ where the solution to the differential equations is defined.
%
We restrict our attention to a specific definition of these dynamical systems:
%
\begin{definition}
A discrete-time nonlinear system is denoted as
\begin{equation}
  x^{+} = f(x)
\label{eq:sys}
\end{equation}
where $f: \reals^{n} \rightarrow \reals^{n}$ is a nonlinear function.
\end{definition}
%
 Intuitively, the function $f$ takes input a state of the system and outputs the next step of the system evolved according to the non-linear dynamics.
%
Here, the function $f$ generally represents some discretized version of some specified continuous non-linear dynamical systems. Recall that a dynamical system is considered \emph{linear} if its dynamics can be expressed as

$$
x' = Ax, \quad A \in \reals^{n \times n}
$$
%
Otherwise, we deem the system to be \emph{nonlinear}.
%
Hence, in particular, the function $f$ cannot be expressed as some matrix $A \in \reals^{n \times n}$.

Examples of prominent non-linear dynamical systems include the Lotka-Volterra predator-prey model \cite{wangersky1978lotka}, FitzHugh Neuron model \cite{fitzhugh1961impulses}, and recently introduced COVID19 disease model \cite{indiansuper2020supermodel}. Throughout this thesis, we discretize any continuous dynamics through the well-known Euler method. Thus, up to some error term of bounded degree, we can turn any non-linear system into the form given by Equation~\ref{eq:sys}.
%
\begin{example}
The SIR Epidemic model is a 3-dimensional dynamical system governed by the following continuous dynamics:
\begin{align} \label{eqn:sir}
    \begin{split}
        s' &= \beta \cdot s_k i_k \\
        i' &= \beta \cdot s_k i_k - \gamma \cdot i_k \\
        r' &= \gamma \cdot i_k
    \end{split}
\end{align}
where $s,i,r$ represent the fractions of a population of individuals designated as \textit{susceptible}, \textit{infected}, and \textit{recovered} respectively. There are two parameters, namely $\beta$ and $\gamma$, which influence the evolution of the system. $\beta$ is labeled as the contraction rate and $1/\gamma$ is the mean infective period.
%
Discretizing Equation~\ref{eqn:sir} according the Euler method yields the dynamics:
%
\begin{align} \label{eqn:discrete_sir}
    \begin{split}
        s_{k+1} &= s_k - (\beta \cdot s_k i_k)\cdot\Delta \\
        i_{k+1} &= i_k + (\beta \cdot s_k i_k - \gamma \cdot i_k)\cdot\Delta \\
        r_{k+1} &= r_k + (\gamma\cdot i_k)\cdot\Delta
    \end{split}
\end{align}
 Here, $\Delta$ is the discretization step and the index $k \in \mathbb{N}$ simply represents the current step.
%For the benchmarks, we set $\beta = 0.34$, $\gamma=0.05$, and $\Delta=0.5$.
%
Note the non-linear terms $s_ki_k$ which precludes the expression of the dynamics as a linear transformation.
\end{example}

The trajectory of a system that evolves according to Equation~\ref{eq:sys}, denoted as $\xi(x_0)$ is a sequence $x_0, x_1, \ldots $ where $x_{i+1} = f(x_i)$.
%
The $k^{th}$ element in this sequence $x_k$ is denoted as $\xi(x_0,k)$.

\begin{definition}
Given an initial set $\Theta \subseteq \reals^n$, the \emph{reachable set at step $k$}, denoted as $\Theta_k$ is defined as
\begin{equation}
  \Theta_k = \{ \xi(x,k)\: | \: x \in \Theta\}
\label{eq:reachset}
\end{equation}
If we set the number of steps to be some $n \in \mathbb{N}$, we say the \emph{reachable set} is
\begin{equation}
  \Theta = \bigcup_{i=1}^n \Theta_i
\end{equation}
\end{definition}

%
We will see in a future section an example of the reachable set of the discretized SIR model presented in Equation~\ref{eqn:discrete_sir}.

\section{Parallelotope-based Reachability}

\subsection{Parallelotopes}
\label{sec:parallelotope}
A parallelotope $P$ is a set of states in $\mathbb{R}^{n}$ captured by the tuple $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ is a matrix and $ç$ is a column vector. We impose the condition that $\Lambda_{i+n} = -\Lambda_{i}$ for all $i \in \{1, \ldots, n\}$ such that
%
\begin{equation}
\label{eq:halfspace}
x \in P \mbox{ if and only if } \Lambda x \leq c.
\end{equation}
We deem $\Lambda$ as the \emph{template direction matrix} where $\Lambda_i$ denotes the $i^{th}$ row of $\Lambda$ called the \emph{$i^{th}$ template direction}. The column vector $c$ is called the \emph{offset vector} with $c_i$ denoting the $i^{th}$ element of $c$.
%
If we unpack Equation~\ref{eq:halfspace}, we can re-express the inequalities as a conjection of half-space constraints. If we define $c_{u} = [c_1, c_2, \cdots, c_n]^T$ and $c_{l} = [c_{n+1}, c_{n+2}, \cdots, c_{2n}]^T$, then Equation~\ref{eq:halfspace} tells us that:
%
\begin{align*}
  & \Lambda_i x \leq (c_u)_i \\
  -& \Lambda_i x \leq (c_l)_i
\end{align*}
%
Combining these two observation yields the \emph{half-space representation} of parallelotope $P$.
%
\begin{definition}
The half-space representation of parallelotope $P$ is tuple  $\langle \Lambda, c\rangle$ where $\Lambda \in \mathbb{R}^{2n \times n}$ and $\Lambda \in \mathbb{R}^{2n \times n}$ such that if $\overline{\Lambda}$ represents the submatrix of $\Lambda$ consisting of the first $n$ rows of $\Lambda$, then
\begin{equation}
\label{eq:halfspaceconst}
P =\{x ~|~ -c_l \leq \overline{\Lambda} x \leq c_u\}
\end{equation}
\end{definition}
%
\noindent Alternatively, a parallelotope can also be represented in a \emph{generator representation}
%
\begin{definition}
The generator representation of a parallelotope $P$ is a tuple of vectors $\langle v, g_1, \ldots, g_n\rangle$ such that $v,g_1, \cdots g_n \in \reals^n$. The vector  $v \in \mathbb{R}^n$ is called the \emph{anchor} and the $g_i \in \mathbb{R}^n$, are called the \emph{generators}. The parallelotope is defined as the set:
$$
P := \{ x ~|~ \exists \alpha_1, \ldots, \alpha_n \in [0,1], \; x = v + \sum_{i=1}^n \alpha_i g_i \}
$$
\end{definition}
%
\noindent This representation is very similar to Zonotopes~\cite{girard2005reachability,althoff2010computing} and Star sets~\cite{duggirala2016parsimonious}.
%
In particular, a parallelotope is a special case of a zonotope where the number of generators is exactly the dimension of the system $n$.

There is a simple method to convert from the half-space representation of $P$ to its generator representation.
%
Obtain vertex $v_1$ by solving the linear equation $\Lambda x = c_{l}$.
%
The $j+1$ vertex is obtained by solving the linear equation $\Lambda x = \mu_{j}$ where $\mu_{j}[i] = c_{l}[i]$ when $i \neq j$ and $\mu_{j}[j] = c_{u}[j]$.
%
The anchor $v$ of the parallelotope is the vertex $v_1$ and the generators will be $g_{i} = v_{i+1} - v_{1}$. Refer to \cite{dang2014parameter} for a more detailed exposition of this procedure.

%
As a final remark, notice that for a parallelotope $P$, the generator representation also defines an affine transformation that maps $[0,1]^{n}$ to $P$.
%
Let us denote this affine transformation associated to $P$ as $T_P$.

\begin{example}
\label{ex:simple_ptope}
Consider the 2D plane, namely $\reals^2$. We can construct a couple of simple examples of parallelotopes. First, if we define our parallelotope's template direction matrix to be
%come back to this
\end{example}

\begin{example}
\label{ex:ptope}
Once again, consider the space $\reals^2$ and the parallelotope $P$ given in half-plane representation as $0 \leq x-y \leq 1$, $0 \leq y \leq 1$.
%
This is a parallelotope with vertices at $(0,0)$, $(1,0)$, $(2,1)$, and $(1,1)$.
%
In the half-space representation, the template directions of the parallelotope $P$ are given by the directions $[1, -1]$ and $[0, 1]$.
%
The half-space representation in matrix form is given as follows:

\begin{equation}
  \begin{bmatrix} 0 \\ 0 \end{bmatrix} \leq \begin{bmatrix}  1 & -1 \\ 0 &  1 \end{bmatrix}  \begin{bmatrix} x \\ y \end{bmatrix} \leq \begin{bmatrix} 1 \\ 1 \end{bmatrix}. \label{eq:ptopeexample}
\end{equation}

To compute the generator representation of $P$, we need to compute the \emph{anchor} and the \emph{generators}.
%
The anchor is obtained by solving the linear equations $x-y = 0, y = 0$.
%
Therefore, the anchor $a$ is the vertex at origin $(0,0)$
%
To compute the two generators of the parallelotope, we compute two vertices of the parallelotope.
%
Vertex $v_1$ is obtained by solving the linear equations $x - y = 1, y = 0$.
%
Therefore, vertex $v_1$ is the vertex $(1,0)$.
%
Similarly, vertex $v_2$ is obtained by solving the linear equations $x-y = 0, y = 1$.
%
Therefore, $v_2$ is the vertex $(1,1)$.
%
The generator $g_1$ is the vector $v_1 - a$, that is $(1,0)- (0,0) = (1,0)$
%
The generator $g_2$ is the vector $v_2 - a$, that is $(1,1) - (0,0) = (1,1)$.
%
Therefore, all the points in the paralellotope can be written as $(x,y) = (0,0) + \alpha_1 (1,0) + \alpha_2(1,1)$, $\alpha_1, \alpha_2 \in [0,1]$.
\end{example}
%
%We denote this affine transformation as $T_{p}$.
%
\begin{definition}
A parallelotope bundle $Q$ is a set of parallelotopes $\{P_1, \ldots, P_m\}$ such that
  $$ Q = \bigcap_{i=1}^{m}P_i $$.
\end{definition}
%
\noindent This parallelotope bundle will be the geometric data structure which will enclose the region we compute to be the approximation of the exact reachable set.

\subsection{Bernstein Polynomials}
\label{sec:bernstein}
In this section, we define Bernstein polynomials and state some of their enclosure properties. A \emph{multi-index} ${\bf i}$ of length $n$ is defined as tuple of $n$ elements ${\bf i} = (i_1,\cdots, i_n)$ such that each $i_k \in \mathbb{N}$. Furthermore, we order the multi-indices as follows: if ${\bf i}$ and ${\bf j}$ are two multi-indices of length $n$, then
\[
  {\bf i} \leq {\bf j} \iff i_k \leq j_k, \quad 1 \leq  k \leq n
\]
Finally, we generalize the product of binomial coefficients as:
\[
  \binom{\boldi}{\boldj} := \prod_{k=1}^n \binom{i_k}{j_k}
\]
%
Given two multi-indices $\boldi$ and $\boldd$ of size $n$, where $\boldi \leq \boldd$, the Bernstein basis polynomial of degree $\boldd$ and index $\boldi$ is defined as:
\begin{equation}
\label{eq:bernstein_basis_poly}
\mathcal{B}_{(\boldi,\boldd)}(\boldx) = \beta_{i_1,d_1}(x_1) \beta_{i_2,d_2}(x_2)\ldots \beta_{i_n,d_n}(x_n).
\end{equation}
%
where for $i,d,x \in \reals$:
%
\begin{equation}
\beta_{i,d}(x) = \binom{d}{i}x^{i}(1-x)^{d - i}
\end{equation}
%
Let $p: \reals^n \rightarrow \reals$ of be a real polynomial of degree at most $\boldd$. We can express $p$ as a linear combination of monomials of degree at most $\boldd$:
\[
p(\boldx) = \sum_{\boldi \leq \boldd} a_\boldi \cdot \boldx^\boldi
\]
where $\boldx^\boldi$ represents the monomial $x_1^{i_1}x_2^{i_2}\cdots x_n^{i_n}$. Every such real polynomial $p$ can be represented as linear combination of Bernstein basis polynomials of of degree $\boldd$:
%
\begin{equation}
  p(\boldx) = \sum_{\boldi \leq \boldd} b_{\boldi} \cdot \mathcal{B}_{(\boldi,\boldd)}(\boldx)
\end{equation}
where $b_\boldi$ denotes the \emph{$\boldi^{th}$ Bernstein Coefficient}:

\begin{equation}
b_{\boldi} = \sum_{\boldj \leq \boldi} \frac{\binom{\boldi}{\boldj}}{\binom{\boldd}{\boldj}}\cdot a_{\boldj}
\end{equation}
%
\noindent In other words, given a polynomial $p(x_1,\ldots,x_n) = \sum_{j \in J} a_j {\bf x}_{j}$ where $J$ is a set of multi-indices iterating through the degrees found in $p$ with $a_j \in \mathbb{R}$, then $p(x_1,\ldots,x_n)$ can be converted into its counterpart under the Bernstein basis, $p(x_1,\ldots,x_n) = \sum_{j \in J} b_j \mathcal{B}_j $ where $b_j$ are the corresponding Bernstein coefficients.

The primary advantage of the Bernstein representation of a polynomial $p(x_1,...,x_n)$ is that an upper bound on the supremum and lower bound on the infimum of $p(x_1,...,x_n)$ in $[0,1]^{n}$ can be computed purely by observing the coefficients of the polynomial in the Bernstein basis. Specifically, the upper and lower bounds of $p(x_1,\ldots,x_n)$ over $[0,1]^n$ are bounded by the Bernstein coefficients:
$$
\min \{b_1, \ldots, b_m\} ~~\leq~~ \inf_{x \in [0,1]^n} p(x) ~~~~~\leq~~~~ \sup_{x \in [0,1]^{n}} p(x) ~~\leq~~ \max \{b, \ldots, b_m\}.
$$

As mentioned earlier, a parallelotope $P$ can also be represented as an affine transformation $T_{p}$ from $[0,1]^{n}$ to $P$.
%
Therefore, upper bounds on the suprenum of a polynomial function $p$ over $P$ is equivalent to upper bound of $p \circ T_{p}$ over $[0,1]^{n}$.
%
A similar argument follows for the lower bound on the infimum.
%
The crux of the reachability algorithm involves exploiting this property of Bernstein polynomials to approximate the solution of certain non-linear optimization problem invovilng polynomial predicates over the unitbox, $[0,1]^n$.
%
We will cover this algorithm in the upcoming section.
%
For a more rigorous exposition on Bernstein polynomials and their properties, refer to \cite{garloff2003bernstein}.

\input{staticalg}