### 2022-07-03 thesis/formats.tex
More progress on static algorithm. Starrting on dynamic part

\chapter{Dynamic Paralleotope Bundles}

\subsection{Local Linear Approximations}
\label{sec:lin_app}

\subsection{Principal Component Analysis}
\label{sec:pca}

\subsection{The Dynamic Algorithm}
\label{sec:dyna_algo}