### 2022-07-05 thesis/conclusion.tex
Adding evaluation tables and plots

\chapter{Conclusion}
\label{chap:conclusion}
In this paper, we investigated two techniques for generating templates dynamically: first using linear approximation of the dynamics, and second using PCA.
%
We demonstrated that these techniques improve the accuracy of reachable set by an order of magnitude when compared to static or random template directions.
%
We also observed that both these techniques improve the accuracy of the reachable sets for different benchmarks.
%
In future, we intend to investigate Koopman linearization techniques for computing alternative linear approximation template directions \cite{bak2021reachability}.
%
We also wish to investigate the use of a massively-parallel implementation using HPC hardware such as GPUs for optimizing over an extremely large number of parallelotopes and their template directions. This is inspired by the approach behind the recent tool \emph{PIRK} \cite{devonport2020pirk}.