### 2022-07-28 thesis/bern_prune_deg_output.tex
Draft of RT reachability written

\begin{verbatim}
            -----------------------------------
            Computing Step 0
            ------------------------------------

            BASE VERTEX: [0.005, 2.0]
            GENERATORS: [[-0.004, 0.0], [0.0, -0.004999999999999893]]

            ------------------------------------
            Stats for Template Direction: [1. 0.]
            ------------------------------------

            Chosen Max Monom Deg: (0, 1),  Chosen Min Monom Deg: (0, 0)

            Actual Max Monom Deg: (0, 0), Actual Min Monom Def: (1, 1)

            Actual Bernstein degree: [1, 1]

            Actual Functional Comp Poly: -0.004*x - 0.000499999999999989*y + 0.205
            Returned set of Maximum Monom from Lookup Table:

            [((0, 1), 0.204500000000000),
             ((1, 0), 0.201000000000000),
             ((1, 1), 0.200500000000000)]

            Returned set of Minimum Monom from Lookup Table:
            [((0, 0), 0.205000000000000)]

            Returned set of ALL Actual Monomials and their ALL Computed Acutal Bernstein Coeff:
            [((0, 0), 0.20500000000000002),
             ((0, 1), 0.20450000000000002),
             ((1, 0), 0.201),
             ((1, 1), 0.2005)]

            Difference between Queried Max Coeff and Actual Max Coeff: -0.000500000000000000

            Difference between Queried Min Coeff and Actual Min Coeff: 0.00450000000000000

            ------------------------------------
            Stats for Template Direction: [0. 1.]
            ------------------------------------

            Chosen Max Monom Deg: (3, 3),  Chosen Min Monom Deg: (0, 0)

            Actual Max Monom Deg: (2, 0), Actual Min Monom Def: (0, 1)

            Actual Bernstein degree: [2, 1]

            Actual Functional Comp Poly: 0.0004*x - 0.00499999999999989*y + 0.1*(1 - 2.5e-5*(1 - 0.8*x)**2)*(2.0 - 0.00499999999999989*y) + 1.9995
            Returned set of Maximum Monom from Lookup Table:

            [((3, 3), 2.18359741950000)]

            Returned set of Minimum Monom from Lookup Table:
            [((0, 0), 2.19949500000000)]

            Returned set of ALL Actual Monomials and their ALL Computed Acutal Bernstein Coeff:
            [((0, 0), 2.199495),
             ((0, 1), 2.1939950125000003),
             ((1, 0), 2.1996990000000003),
             ((1, 1), 2.1941990025000004),
             ((2, 0), 2.1998998000000003),
             ((2, 1), 2.1943998005000007)]

            Difference between Queried Max Coeff and Actual Max Coeff: -0.0163023804999995

            Difference between Queried Min Coeff and Actual Min Coeff: 0.00549998749999991
  \end{verbatim}