### 2022-07-28 thesis/bern_prune_wrapping_error.tex

\begin{verbatim}
------------------------------------
Computing Step 3
------------------------------------

------------------------------------
Stats for Template Direction: [1. 0.]
------------------------------------

Chosen Max Monom Deg: (1, 1),  Chosen Min Monom Deg: (0, 0)

Actual Max Monom Deg: (1, 1), Actual Min Monom Deg: (0, 0)

Actual Bernstein degree: [1, 1]

Actual Sympy Poly: 0.007306445824668*x + 0.0111954493962119*y + 0.906973678980699
Returned set of Maximum Monom from lookup_table:

[((1, 1), 0.925475574201579)]

Returned set of Minimum Monom from lookup_table:
[((0, 0), 0.906973678980699)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 0.906973678980699),
 ((0, 1), 0.9181691283769109),
 ((1, 0), 0.914280124805367),
 ((1, 1), 0.9254755742015789)]

Difference between Queried Max Coeff and Actual Max Coeff: 0

Difference between Queried Min Coeff and Actual Min Coeff: 0

------------------------------------
Stats for Template Direction: [0. 1.]
------------------------------------

Chosen Max Monom Deg: (0, 3),  Chosen Min Monom Deg: (0, 0)

Actual Max Monom Deg: (0, 1), Actual Min Monom Def: (2, 0)

Actual Bernstein degree: [2, 1]

Actual Sympy Poly: -0.0007306445824668*x + 0.111954493962119*y + 0.1*(1 - 0.435728408759507*(0.0110687411271527*x + 1)**2)*(0.111954493962119*y + 2.46876406785527) + 2.40275434063576

Returned set of Maximum Monom from lookup_table:

[((1, 2), 2.77693923047705),
 ((2, 1), 2.65709768697245),
 ((3, 1), 2.65546013926846),
 ((0, 2), 2.77860321954750),
 ((2, 2), 2.77526086680021),
 ((3, 2), 2.77356812851698),
 ((1, 3), 2.89515700322449),
 ((1, 1), 2.65872145772961),
 ((0, 3), 2.89687498755508),
 ((2, 0), 2.53893450714470),
 ((3, 0), 2.53735215001995),
 ((2, 3), 2.89342404662796),
 ((3, 3), 2.89167611776549)]

Returned set of Minimum Monom from lookup_table:
[((0, 0), 2.54205968353236)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 2.542059683532361),
 ((0, 1), 2.6603314515399323),
 ((1, 0), 2.5405036849821685),
 ((1, 1), 2.658721457729609),
 ((2, 0), 2.5389345071447),
 ((2, 1), 2.657097686972454)]

Difference between Queried Max Coeff and Actual Max Coeff: 0.236543536015143

Difference between Queried Min Coeff and Actual Min Coeff: 0.00312517638766119

------------------------------------
Computing Step 4
------------------------------------

------------------------------------
Stats for Template Direction: [1. 0.]
------------------------------------

Chosen Max Monom Deg: (1, 1),  Chosen Min Monom Deg: (0, 0)\ n
Actual Max Monom Deg: (0, 0), Actual Min Monom Def: (1, 1)

Actual Bernstein degree: [1, 1]

Actual Sympy Poly: -0.0185018952208799*x - 0.0354815304022714*y + 1.21516307295709
Returned set of Maximum Monom from lookup_table:

[((1, 1), 1.16117964733394)]

Returned set of Minimum Monom from lookup_table:
[((0, 0), 1.21516307295709)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 1.2151630729570864),
 ((0, 1), 1.1796815425548148),
 ((1, 0), 1.1966611777362064),
 ((1, 1), 1.161179647333935)]

Difference between Queried Max Coeff and Actual Max Coeff: -0.0539834256231513

Difference between Queried Min Coeff and Actual Min Coeff: 0.0539834256231513

------------------------------------
Stats for Template Direction: [0. 1.]
------------------------------------

Chosen Max Monom Deg: (3, 0),  Chosen Min Monom Deg: (0, 3)\ n
Actual Max Monom Deg: (2, 0), Actual Min Monom Def: (0, 1)

Actual Bernstein degree: [2, 1]

Actual Sympy Poly: 0.00185018952208799*x - 0.354815304022714*y + 0.1*(1 - 0.856505038443742*(1 - 0.0199917704331005*x)**2)*(2.89687498755508 - 0.354815304022714*y) + 2.80432743013492

Returned set of Maximum Monom from lookup_table:

[((0, 1), 2.48598940176842),
 ((1, 2), 2.13075300161303),
 ((2, 1), 2.49645813555339),
 ((3, 1), 2.50156197271691),
 ((0, 2), 2.12608267690468),
 ((2, 2), 2.13534845254406),
 ((1, 0), 2.85178155552809),
 ((3, 2), 2.13986902969777),
 ((1, 3), 1.77023872465550),
 ((1, 1), 2.49126727857056),
 ((0, 3), 1.76617595204093),
 ((2, 0), 2.85756781856273),
 ((3, 0), 2.86325491573606),
 ((2, 3), 1.77423876953473),
 ((3, 3), 1.77817608667862)]

Returned set of Minimum Monom from lookup_table:
[((0, 1), 2.48598940176842),
 ((3, 1), 2.50156197271691),
 ((2, 2), 2.13534845254406),
 ((1, 0), 2.85178155552809),
 ((3, 2), 2.13986902969777),
 ((1, 3), 1.77023872465550),
 ((0, 0), 2.84589612663217),
 ((0, 3), 1.76617595204093),
 ((3, 0), 2.86325491573606),
 ((2, 3), 1.77423876953473),
 ((3, 3), 1.77817608667862)]

Returned set of ALL Monomials and their ALL Computed Actual Bernstein Coeff:
[((0, 0), 2.8458961266321676),
 ((0, 1), 2.485989401768422),
 ((1, 0), 2.8517815555280945),
 ((1, 1), 2.491267278570564),
 ((2, 0), 2.8575678185627256),
 ((2, 1), 2.496458135553395)]

Difference between Queried Max Coeff and Actual Max Coeff: 0.00568709717333471

Difference between Queried Min Coeff and Actual Min Coeff: -0.719813449727491
\end{verbatim}