### 2022-07-30 thesis/bern_prune_deg_output.tex
Second draft of thesis finished

\begin{figure}
\begin{verbatim}
    -----------------------------------
              Computing Step 0
    -----------------------------------

    BASE VERTEX: [0.005, 2.0]
    GENERATORS: [[-0.004, 0.0], [0.0, -0.0049999999999999]]
    ....
    ....
    -------------------------------------
    Stats for Template Direction: [0. 1.]
    -------------------------------------

    Queried Max Monom Deg: (3, 3)
    Queried Min Monom Deg: (0, 0)

    Actual Max Monom Deg: (2, 0)
    Actual Min Monom Def: (0, 1)

    Actual Composed Poly:
    0.0004*x - 0.00499999999999989*y
    + 0.1*(1 - 2.5e-5*(1 - 0.8*x)**2)*(2.0 - 0.00499999999999989*y)
    + 1.9995

    Total Poly Degree: [2, 1]

    Returned set of Maximum Monom from Lookup Table:
    [((3, 3), 2.18359741950000)]

    Returned set of Minimum Monom from Lookup Table:
    [((0, 0), 2.19949500000000)]

    Set of ALL basis degrees and true Bernstein coeffs:
    [((0, 0), 2.199495),
     ((0, 1), 2.1939950125000003),
     ((1, 0), 2.1996990000000003),
     ((1, 1), 2.1941990025000004),
     ((2, 0), 2.1998998000000003),
     ((2, 1), 2.1943998005000007)]

    Difference between Queried Max Coeff and Actual Max Coeff:
    -0.0163023804999995

    Difference between Queried Min Coeff and Actual Min Coeff:
    0.00549998749999991
  \end{verbatim}
  \caption{Output for First step of Reachability algorithm with the Vanderpol Model.}
  \label{fig:bern_prune_deg_output}
\end{figure}