### 2022-07-30 thesis/bern_prune_wrapping_error.tex

\begin{figure}
\begin{verbatim}
------------------------------------
Computing Step 4
------------------------------------

------------------------------------
Stats for Template Direction: [0. 1.]
------------------------------------

Chosen Max Monom Deg: (3, 0)
Chosen Min Monom Deg: (0, 3)
...
Returned set of Maximum Monom from lookup_table:
[((0, 1), 2.48598940176842),
 ((1, 2), 2.13075300161303),
 ((2, 1), 2.49645813555339),
 ((3, 1), 2.50156197271691),
 ((0, 2), 2.12608267690468),
 ((2, 2), 2.13534845254406),
 ((1, 0), 2.85178155552809),
 ((3, 2), 2.13986902969777),
 ((1, 3), 1.77023872465550),
 ((1, 1), 2.49126727857056),
 ((0, 3), 1.76617595204093),
 ((2, 0), 2.85756781856273),
 ((3, 0), 2.86325491573606),
 ((2, 3), 1.77423876953473),
 ((3, 3), 1.77817608667862)]

Returned set of Minimum Monom from lookup_table:
[((0, 1), 2.48598940176842),
 ((3, 1), 2.50156197271691),
 ((2, 2), 2.13534845254406),
 ((1, 0), 2.85178155552809),
 ((3, 2), 2.13986902969777),
 ((1, 3), 1.77023872465550),
 ((0, 0), 2.84589612663217),
 ((0, 3), 1.76617595204093),
 ((3, 0), 2.86325491573606),
 ((2, 3), 1.77423876953473),
 ((3, 3), 1.77817608667862)]

Difference between Queried Max Coeff and Actual Max Coeff:
0.00568709717333471

Difference between Queried Min Coeff and Actual Min Coeff: -
0.719813449727491
\end{verbatim}

\caption{Example of Wrapping Error Obstacle demonstrated for Step 4 for Vanderpol Model.}
\label{fig:bern_prune_wrapping_output}
\end{figure}